---
title: "7PAVITPR Introduction to Statistical Programming \nAssignment - R Part"
author: "Yui Hang, Wong (Harris)\nK21115599"
date: "Compiled: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
---

# Library
```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(haven))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(knitr))
```

\newpage
# 1 Manipulating dataframes

## 1.1
```{r}
file <- here("data", "source", "ESS.sav")
ess_raw <- haven::read_sav(file)
```

## 1.2
```{r}
ess <- ess_raw %>% select(
        idno, cntry, stfeco, gincdif, advbach, gndr, yrbrn,
        wkhtot, jbscr, mainact, hltherb
)

head(ess)
```

## 1.3
```{r}
ess <- ess %>% rename(
        idno = idno,
        country = cntry,
        satisfied_econ = stfeco,
        reduce_diff = gincdif,
        first_if_backache = advbach,
        gender = gndr,
        yob = yrbrn,
        working_hours = wkhtot,
        job_secure = jbscr,
        main_activity = mainact,
        herbal_remedies = hltherb
)
```

## 1.4
```{r}
year_of_collection <- 2004
ess <- ess %>% mutate(age = year_of_collection - yob) 

ess$age[1:10]
```

## 1.5
```{r}
f_list = list(
        "sd" = sd,
        "mean" = mean,
        "oldest" = max,
        "youngest" = min
)

ess %>% summarise(across(age, f_list, na.rm = T))
```

## 1.6
```{r}
p1.6.1 <- ess %>% 
        drop_na(age) %>% 
        ggplot(aes(x = age)) +
        geom_histogram(bins = 30)

p1.6.2 <- p1.6.1 + facet_wrap(~country)

p1.6.1
p1.6.2
```

## 1.7
```{r}
#! NA treatment?
ess <- ess %>% mutate(female = as.numeric(gender == 2))
```

## 1.8
```{r}
ess <- ess %>% 
        mutate(work_over_50h = working_hours  > 50)
```

## 1.9

Tabulate
```{r}
counts <- ess %>% count(work_over_50h)
counts
```

(i) Overall proportion of people working long hours
```{r}
counts %>% 
        mutate(prop = n/sum(n)) %>% 
        filter(work_over_50h) %>% 
        pull(prop)
```

Proportion by country of people working long hours
```{r}
ess %>% 
        group_by(country) %>% 
        count(work_over_50h) %>% 
        mutate(prop = n / sum(n)) %>% 
        ungroup() %>% 
        filter(work_over_50h) %>% 
        select(country, prop) %>% 
        kable()
```

## 1.10
```{r}
ess <- ess %>%
        mutate(first_if_backache = as.numeric(first_if_backache %in% 4:5))
```

## 1.11
```{r}
bycountry <- ess %>% 
        group_by(country) %>% 
        count(first_if_backache) %>% 
        mutate(prop = n / sum(n)) %>% 
        filter(first_if_backache == 1) %>% 
        select(country, prop) %>% 
        rename(seekdoctor = prop) %>% 
        ungroup()

head(bycountry)
```

## 1.12
```{r}
load(here("data", "source", "health.Rdata"))

health_long <- health %>%
        pivot_longer(cols = -measure,
                     names_to = "cntry") %>%
        tidyr::separate(measure,
                        into = c(NA, "feature", "year"), # drop the first name
                        sep = "_") %>%
        pivot_wider(names_from = "feature",
                    values_from = "value") %>%
        rename(healthspend = pcexhe, doctors = phyrape)

head(health_long)
```

## 1.13
```{r}
year_of_collection <- 2004

merged <- health_long %>% 
        filter(year == year_of_collection) %>% # filter for the year of collection only
        select(-year) %>% 
        full_join(bycountry, by = c("cntry" = "country"))

kable(merged)

merged %>% 
        transmute(
                only_left = !is.na(healthspend) & !is.na(doctors),
                only_right = !is.na(seekdoctor),
                both = only_left & only_right
        ) %>% 
        summarise(across(everything(), sum))



```

## 1.14
```{r}
p1.14 <- merged %>% 
        ggplot(aes(x = healthspend, y = seekdoctor)) +
        geom_point() +
        ggplot2::stat_smooth(method = "lm", se = F) +
        labs(title = "National expenditure on healthcare per capita \nagainst the proportion of people who report going to a doctor/nurse ﬁirst \n in the event they experience serious backache in European countries",
             x = "National expenditure on healthcare per capita",
             y = "The proportion of people who report going to a doctor/nurse first \nin the event they experience serious backache ")
p1.14
```


# 2 Random Numbers, Skewness and Kurtosis

## set seed
```{r}
set.seed(0)
```

## 2.1
```{r}
n.seq <- c(100, 200, 10^3, 10^4, 10^5)
```

## 2.2
```{r}
samples <- n.seq %>% 
        map(rnorm, mean = 0, sd = 1)
```

A list can be used to store vectors of different lengths.

## 2.3

Helper plotting functions
```{r}
#' Plot a histogram of distribution from a numeric vector using ggplot
plot_histo <- function(x, title = NULL) {
        p <- x %>% 
                as.data.frame() %>% 
                ggplot(aes(x = x)) +
                geom_histogram(bins = 30) +
                labs(title = title)
        return(p)
}

#' Call plot_histo to produce a list of ggplot objects, 
#' then wrap the objects into a 1-row panel using patchwork
plot_histo_panel <- function(distributions, titles) {
        plot_list <- map2(
                .x = distributions,
                .y = titles,
                .f = plot_histo
        ) 
        pw <- plot_list %>% patchwork::wrap_plots(nrow = 1)
        return(pw)
}
```

```{r}
titles <- as.character(n.seq)
p2.3 <- plot_histo_panel(samples, titles)
p2.3
```

## 2.4

The histograms show the Central Limit Theorem, in that as the sample size increases and approaches infinity, the distribution () of
#! central limit theorem? The law of large numbers? How to describe?

## 2.5
```{r}
bimodal <- c(
        rnorm(mean = 0, sd = 1, n = 10000),
        rnorm(mean = 4, sd = 1, n = 10000)
        )
p2.5 <- plot_histo(bimodal)
p2.5
```

## 2.6
```{r}
trimodal <- c(
        rnorm(mean = -4, sd = 1, n = 10000),
        rnorm(mean = 0,  sd = 1, n = 10000),
        rnorm(mean = 4,  sd = 1, n = 10000)
        )
p2.6 <- plot_histo(trimodal)
p2.6
```

## 2.7
```{r}
right_skewed <- c(
        rnorm(mean = 0, sd = 1, n = 1000),
        rnorm(mean = 2, sd = 2, n = 1000)
)
left_skewed <- c(
        rnorm(mean = 2, sd = 2, n = 1000),
        rnorm(mean = 0, sd = 1, n = 1000)
)

skewed_distributions <- list(
        "Left-skewed" = left_skewed, 
        "Right-skewed" = right_skewed
)

titles <- names(skewed_distributions)
plot_histo_panel(skewed_distributions, titles)

skewed_distributions %>% 
        map_dbl(e1071::skewness)
```

## 2.8
```{r}
leptokurtic <- c(
        rnorm(10000, mean = 0, sd = 1),
        rnorm(100, mean = 5,  sd = 1),
        rnorm(100, mean = 10, sd = 1)
)
platykurtic <- c(
        rnorm(10000, mean = 0, sd = 1),
        rnorm(10000, mean = 0, sd = 2)
        )

kurtosis_distributions <- list(
        "Leptokurtotic" = leptokurtic, 
        "Platykurtic"   = platykurtic
)

titles <- kurtosis_distributions %>% 
        map(e1071::kurtosis) %>% 
        map_chr(function(x) paste("Kurtosis:", x))

p2.8 <- plot_histo_panel(kurtosis_distributions, titles)
p2.8
```





